Increased perceptions of autonomy through choice fail to enhance motor skill retention

### 2021-11-03 docs/r1_submission/r1_manuscript.Rmd
update: begin moving revisions from gdocs to rmd

Over the years, this manipulation has been described using a variety of names (e.g., self- and/or learner-controlled; -regulated; -directed; -selected), but more recently it has been subsumed by autonomy-support. In fact, autonomy-support is considered such a robust learning variable that it is one of three key pillars in the recently proposed "OPTIMAL" theory of motor learning [@wulf2016]. @wulf2016 argued that providing learners with opportunities for choice---considered an autonomy-supportive practice manipulation---can facilitate motor performance and learning by enhancing learner's expectancies for success, and by allowing the learner to maintain their attentional focus on the task by reducing the need for self-regulatory activity [e.g., @abdollahipour2017; @chua2018; @wulf2018]. In other words, these psychological and attentional benefits, and concomitant increases in performance and learning are a by-product of exercising choice during practice.

Despite its prominent role within the "OPTIMAL" theory of motor learning, there are reasons to doubt the importance of autonomy-support through the provision of choice during practice. First, if the benefits are primarily the result of having opportunities for choice then learning differences should not emerge between different self-controlled groups. However, in experiments where different groups of participants have choice over their feedback schedule, such learning differences have been found when this choice is made *after* rather than before a performance attempt [@chiviacowsky2005; @carter2014], when different criteria for success are provided to participants [@chiviacowsky2012], and when the absolute number of feedback choice opportunities are limited compared to unlimited at the outset of practice [@hansen2011; @patterson2011]. Second, the notion that practicing in a self-controlled group is in fact more autonomy-supportive than practicing in a yoked group has predominantly been assumed rather than supported empirically.

>>>>

Over the years, this manipulation has been described using a variety of names (e.g., self- and/or learner-controlled; -regulated; -directed; -selected), but more recently it has been subsumed by autonomy-support. In fact, autonomy-support is considered such a robust learning variable that it is one of three key pillars in the recently proposed "OPTIMAL" theory of motor learning [@wulf2016]. @wulf2016 argued that providing learners with opportunities for choice---considered an autonomy-supportive practice manipulation---can facilitate motor performance and learning through enhanced expectancies for success and having the learner maintain their attentional focus on the task thereby reducing the need for self-regulatory activity [e.g., @abdollahipour2017; @chua2018; @wulf2018]. In other words, these psychological and attentional benefits, and concomitant increases in performance and learning are a by-product of exercising choice during practice. Overall, @wulf2016's "OPTIMAL" theory of motor learning provides a motivational account for the learning advantage of self-controlled practice conditions over yoked practice conditions. [ADD A FOOTNOTE ABOUT INFORMATION-PROCESSING BUT THAT IT IS THE MOTIVATIONAL VIEW THAT SEEMS TO HAVE GARNERED THE MOST ATTENTION]

Despite its prominent role within the "OPTIMAL" theory of motor learning, there are reasons to doubt the importance of autonomy-support through the provision of choice during practice. First, if the learning benefits arise out of the opportunity for choice itself then learning differences should not emerge between different self-controlled groups. However, in experiments where different groups of participants have choice over their feedback schedule, such learning differences have been found when this choice is made *after* rather than before a performance attempt [@chiviacowsky2005; @carter2014], when different criteria for success are provided to participants [@chiviacowsky2012], and when the absolute number of feedback choice opportunities are limited compared to unlimited at the outset of practice [@hansen2011; @patterson2011]. Second, the notion that practicing in a self-controlled group is in fact more autonomy-supportive than practicing in a yoked group has predominantly been assumed rather than supported empirically.
